Wikipedia Text Mining and Surprisal Analysis

### 2019-05-14 entropy/childes/non-english_entropies.Rmd
reorganization; working on wiki relative unigram surprisal

  tokens <- sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, stem) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())

>>>>

  tokens <- sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, stem) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    filter(word_order <= sen_length)

### 2019-05-14 surprisal/wiki_surprisal.Rmd
reorganization; working on wiki relative unigram surprisal

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka and Dan Yurovsky
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

@@

```{r imputation and MCA}
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs)
LANGUAGES <- wkpd$Language

FEATURES <- c(1:144)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all WALS features in correct syntax

fair_wf <- wf[rowSums(!is.na(wf)) > 50,] %>%
  filter(language %in% LANGUAGES) %>%
  select(-wals.code, -latitude, -longitude, -glottocode) %>%
  as_tibble() %>%
  clean_names()

mean_wf <- fair_wf

mean_impute <- function(col) {
  counts <- col %>%
    remove_missing() %>%
    unlist() %>%
    table() %>%
    sort(decreasing = T) 
  col[is.na(col),] <- labels(counts[1])
  return(col)
}

for (i in 1:ncol(mean_wf)) {
  mean_wf[, i] <- mean_impute(mean_wf[, i])
}

# Phonology `1A`:`19A`
# Morphology `20A`:`29A`
# Nominal Categories `30A`:`57A`
# Nominal Syntax `58A`:64A`
# Verbal syntax `65A`:`80A`
# Word order `81A`:`97A`, `143A`, `144A`
# Simple clauses `98A`:`120A`
# Complex sentences `121A`:`128A`
# Lexicon `129A`:`138A`
# Sign languages `139A`:`140A`

wa <- fair_wf %>% 
  select(x1a:x10a, language) %>%
  group_by(language) %>%
  slice(1)

wa_langs <- wa %>%
  gather(measure, value, -language) %>%
  group_by(language) %>%
  summarise(value = mean(!is.na(value))) %>%
  arrange(desc(value)) %>%
  filter(value > .6) %>% 
  pull(language)

wa_sub <- wa %>%
  filter(language %in% wa_langs) %>%
  column_to_rownames("language") %>%
  ungroup()
 # mutate_each(~as.numeric(as.factor(.x)))

wa_out <- wa_sub %>%
  MIMCA(ncp = 1)

wa_imputed <- wa_out[["res.MI"]] %>%
  map(~mutate(.x, language = wa_langs)) %>%
  bind_rows(.id = "boot") %>%
  group_by(language, boot) %>%
  gather(measure, value, -language, -boot) %>%
  group_by(language, measure) %>%
  summarise(value = mean(value, na.rm = T)) %>%
  spread(measure, value)
```

@@

```{r hierarchical clustering}
NUM_CLUSTERS <- 10

##read in googlesheet
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

wkpd <- wkpd %>% 
  as.data.frame() %>%
  column_to_rownames(var = "Language")

##do clustering
wk_clust_tree <- wkpd %>% 
              dist() %>%
              hclust("ward.D2")

wk_clust_tree %>% plot(hang = -1)
wk_clusters <- wk_clust_tree %>%
              cutree(k = NUM_CLUSTERS)

wk_clusters %>% 
  t() %>%
  as_data_frame() %>% 
  t() %>%
  as_data_frame(rownames = "language") %>%
  rename(cluster = V1) %>%
  arrange(cluster) %>% 
  View()
```

```{r}
wiki_gs <- gs_title("Wikipedia_relative5")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

FEATURES <- c(1:144)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all WALS features in correct syntax

wf_df <- wf %>%
  as_data_frame() %>%
  filter(language %in% LANGUAGES) %>%
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, `1A`:`144A`) %>%
  group_by(feature) %>%
  mutate(value = factor(value))

```

```{r cogsci glm subset}

wf_many_langs <- wf_df %>%
  group_by(feature) %>%
  summarise(n = sum(!is.na(value))) %>%
  filter(n > 60)

wf_sub_langs <- wf_df %>%
  filter(feature %in% wf_many_langs$feature) %>%
  group_by(language) %>%
  summarise(keep = sum(is.na(value)) == 0) %>%
  filter(keep)

wf_sub <- wf_df %>%
  filter(feature %in% wf_many_langs$feature,
         language %in% wf_sub_langs$language) %>%
  select(language, feature, value) %>%
  mutate(value = as.numeric(as.factor(value)))

wf_pairwise <- expand.grid(language1 = wf_sub$language,
                           language2 = wf_sub$language) %>%
  left_join(wf_sub, by = c("language1" = "language")) %>%
  rename(value1 = value) %>%
  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
  rename(value2 = value) %>%
  mutate(same = value1 == value2) %>%
  select(-value1, -value2) %>%
  left_join(cosines, by = c("language1", "language2"))

models %>%
  mutate(coeffs = map(model, tidy),
         rs = map(model, )) %>%
  select(-data, -model) %>%
  unnest() %>%
  filter(term == "cosine") %>%
  arrange(desc(statistic))
```

>>>>

---
title: Wikipedia Text Mining and Surprisal Analysis
author: Josef Klafka and Dan Yurovsky
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

@@

@@
